---
title: "Library Workshop Exercise"
author: "Your Name"
date: "Type completion date"
output:
  word_document: default
  html_document: default
---
```{r setup, include=FALSE}
# Do not delete or alter this chunk, always keep it in your R Markdown. 
knitr::opts_chunk$set(echo = TRUE)
```

**Go to https://research.library.gsu.edu/c.php?g=972058&p=7685877 and complete the exercise by creating chunks and typing the codes. After completing the exercise, you need to submit an RMD file and a knitted WORD file. I already typed the first two codes as well as the last, but you need to run them. ** 

```{r loading and opening data}
# Run this code to load and open the data
SLID <- read.csv("https://raw.githubusercontent.com/GSU-Robinson/Data/main/SLID.csv")
```

```{r viewing the data using the name}
head(SLID)
```
### Frequency Tables:
Below is the code for the frequency distribution for the variable language in the SLID dataset. 
```{r}
table(SLID$language)
```

### Descriptive Statistics
Below is the code for calculating the descriptive statistics of the variable wages.
```{r}
summary(SLID$wages)
```

We can also calculate the descriptive statistics for all the variables in one command line.
```{r}
summary(SLID)
```

### Cross Tabulation
Below is the code for conducting a crosstabulation and calculating the Chi-squared test.    
```{r}
table(SLID$sex, SLID$language)

chisq.test(table(SLID$sex, SLID$language))
```

### Generating Variables
Below is the code for generating the variable age1 from an already existing variable age. 
```{r}
SLID$age1<-SLID$age
```

### Correlations
Below is the code for calculating the sample Pearson coefficient. 
```{r}
cor(SLID[,c("wages", "age")], use="complete.obs")

cor.test(SLID$wages, SLID$age, method = "pearson")
```

### Linear Regression
Below is the RStudio code for our linear regression. 
```{r}
summary(lm(wages~age + sex + education + language, data= SLID))
```

### Basic Graphing
A pie chart is best for showing the proportions of occurrence of the options in a nominal level variable. 
```{r}
pie(table(SLID$language))
```

```{r, message=FALSE, results=FALSE}
# This is your last chunk
library(car)
scatterplot(wages~education, regLine=FALSE, smooth=FALSE, boxplots=FALSE, data=SLID)
```




