---
title: "R Assignment 2"
author: "W Hunter Giles"
date: '2022-02-21'
output:
  word_document: default
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.  Import the **Demographics** data file (csv format) into a data frame (table) and label it myData. Keep in mind that The R language is case sensitive.

```{r}
# Replace Your with your name and " " with your file location. 

myData <- read.csv("./data/Demographics.csv")

# View(myData)
```

1.  Count the number of males and females in the data.

```{r}
length(which(myData$Sex=='M'))
length(which(myData$Sex=='F'))

```

1.  What percentages of males and females are married?

```{r}
length(which(myData$Sex=='M' & myData$Married=='Y'))/length(which(myData$Sex=='M'))
length(which(myData$Sex=='F' & myData$Married=='Y'))/length(which(myData$Sex=='F'))
```

1.  Of the 10 individuals with the highest income, how many are married males.

```{r}
df2 <- myData[order(-myData$Income),]
df2 <- df2[1:10,]
length(which(df2$Married=='Y' & df2$Sex=='M'))
```

1.  What are the highest and the lowest incomes of males and females?

```{r, message=FALSE}
library(dplyr)
# lowest and highest income for males:
min(filter(myData, myData$Sex=='M')$Income)
max(filter(myData, myData$Sex=='M')$Income)

# lowest and highest income for females
min(filter(myData, myData$Sex=='F')$Income)
max(filter(myData, myData$Sex=='F')$Income)
```

1.  What are the highest and lowest incomes of married and unmarried males?

```{r}
# "lowest and highest income for Unmarried males:"
min(filter(myData, Sex=='M', Married=='N')$Income)
max(filter(myData, Sex=='M', Married=='N')$Income)

# lowest and highest income for married males:
min(filter(myData, Sex=='M', Married=='Y')$Income)
max(filter(myData, Sex=='M', Married=='Y')$Income)
```
